---
title: "Data preprocessing"
output: html_notebook
date: "2024-09-30"
---
# Package setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(survival, JM, lattice, parallel, dplyr, tidyr, readr, writexl, readxl, survminer, psych, pheatmap, rmcorr)
set.seed(123)
```
# Case Study One: MMH, SI

## Read in the short and long forms of data
```{r}
short_data <- read_csv("time_fixed_conservative.csv")
long_data <- read_csv("consv_sds.csv")
```

## Leave two subjects out train-test splits
```{r}
d <- short_data
l <- long_data

treatmentList <- c("Task")
dlevels <- list(
          Task = c("MMH", "SI")
        )

library(combinat)
set.seed(123)
# Number of observations
n <- unique(d$Index)
# Number of combinations to leave two out
p <- 2
# Generate all combinations of leaving two out
combinations <- combinat::combn(n, p)

# List to store train-validation datasets
train_test_datasets <- list()
train_test_sub_datasets <- list()

# Iterate over combinations
for (i in seq(ncol(combinations))) {
  testing_indices <- combinations[, i]
  training_indices <- setdiff(n, testing_indices) # 1:length(n)
  
  # Create train-validation datasets
  training_data_sub <- d[d$Index %in% training_indices,]
  testing_data_sub <- d[d$Index %in% testing_indices,] 
  
  train_test_sub_datasets[[i]] <- list(train_data = training_data_sub, test_data = testing_data_sub)
  
  training_data <- l[l$Index %in% training_indices,]
  testing_data <- l[l$Index %in% testing_indices,]
  
  for (treat in treatmentList) {
    training_data[[treat]] <- factor(training_data[[treat]], level = dlevels[[treat]])
    testing_data[[treat]] <- factor(testing_data[[treat]], level = dlevels[[treat]])
    
    training_data_sub[[treat]] <- factor(training_data_sub[[treat]], level = dlevels[[treat]])
    testing_data_sub[[treat]] <- factor(testing_data_sub[[treat]], level = dlevels[[treat]])
    
  }
  train_test_datasets[[i]] <- list(train_data = training_data, test_data = testing_data)
    train_test_sub_datasets[[i]] <- list(train_data = training_data_sub, test_data = testing_data_sub)
}
```

```{r}
saveRDS(train_test_datasets, file = "lconsvftall_l2so_train_test_datasets.rds")
saveRDS(train_test_sub_datasets, file ="sconsvftall_l2so_train_test_datasets.rds")
```

## Select sensor features that have >=0.2 correlations with fatigue state.
```{r}
# Input the data and lme_form for the current run
rpe_cut <- "consv"

sname <- paste0("s", rpe_cut, "ftall_l2so_train_test_datasets.rds")
lname <- paste0("l", rpe_cut, "ftall_l2so_train_test_datasets.rds")
data_sub_list <- readRDS(sname)
data_list <- readRDS(lname)
```

```{r}
feature_list <- list()
all_features <- c()

for (i in seq_along(data_sub_list))
  {
    data <- data_list[[i]]$train_data

    cor_matrix <- mixedCor(data=data, c=10:72, d=6)$rho
    cfeatures <- data.frame(cor_matrix) |> select (State)
    cfeatures_indices <- apply(cfeatures, 2, function(x) which(abs(x) >= 0.2))
    cfeature_names <- rownames(cfeatures_indices)[-1]

    feature_list[[i]] <- cfeature_names
    all_features <- c(all_features, unlist(cfeature_names))
}

View(all_features)
sort(table(all_features), decreasing = TRUE)
```

# Case Study Two: BP
## Read in the short and long forms of data
```{r}
short_data <- read_csv("short_data4_r0.5.csv")
long_data <- read_csv("long_data4_r0.5.csv")
```

## Leave two subjects out train-test splits
```{r}
d <- short_data
l <- long_data

treatmentList <- c("experimental_condition", "bottle_mass", "picking_rate")
dlevels <- list(
          experimental_condition = c("Condition 1.5-15", "Condition 2.5-10", "Condition 2.5-15", "Condition 2.5-5"),
          bottle_mass = c(1.5, 2.5),
          picking_rate = c(5, 10, 15)
        )

library(combinat)
set.seed(123)
# Number of observations
n <- unique(d$Index)
# Number of combinations to leave two out
p <- 2
# Generate all combinations of leaving two out
combinations <- combinat::combn(n, p)

# List to store train-validation datasets
train_test_datasets <- list()
train_test_sub_datasets <- list()

# Iterate over combinations
for (i in seq(ncol(combinations))) {
  testing_indices <- combinations[, i]
  training_indices <- setdiff(n, testing_indices) # 1:length(n)
  
  # Create train-validation datasets
  training_data_sub <- d[d$Index %in% training_indices,]
  testing_data_sub <- d[d$Index %in% testing_indices,] 
  
  train_test_sub_datasets[[i]] <- list(train_data = training_data_sub, test_data = testing_data_sub)
  
  training_data <- l[l$Index %in% training_indices,]
  testing_data <- l[l$Index %in% testing_indices,]
  
  for (treat in treatmentList) {
    training_data[[treat]] <- factor(training_data[[treat]], level = dlevels[[treat]])
    testing_data[[treat]] <- factor(testing_data[[treat]], level = dlevels[[treat]])
    
    training_data_sub[[treat]] <- factor(training_data_sub[[treat]], level = dlevels[[treat]])
    testing_data_sub[[treat]] <- factor(testing_data_sub[[treat]], level = dlevels[[treat]])
    
  }
  train_test_datasets[[i]] <- list(train_data = training_data, test_data = testing_data)
    train_test_sub_datasets[[i]] <- list(train_data = training_data_sub, test_data = testing_data_sub)
}
```

```{r}
# saveRDS(train_test_datasets, file = "longftr_all_shoulder4_l2so_train_test_datasets.rds")
# saveRDS(train_test_sub_datasets, file ="shortftr_all_shoulder4_l2so_train_test_datasets.rds")
```

## Select sensor features that have >=0.2 correlations with fatigue state.
```{r}
# Input the data and lme_form for the current run
# rpe_cut <- "4"
# 
# data_sub_list <- readRDS("shortftr_all_shoulder4_l2so_train_test_datasets.rds")
# data_list <- readRDS("longftr_all_shoulder4_l2so_train_test_datasets.rds")
```

```{r}
feature_list <- list()
all_features <- c()

for (i in seq_along(data_sub_list))
  {
    data <- data_list[[i]]$train_data
    cor_matrix <- mixedCor(data=data, c=15:68, d=6)$rho
    cfeatures <- data.frame(cor_matrix) |> select (State)
    cfeatures_indices <- apply(cfeatures, 2, function(x) which(abs(x) >= 0.2))
    cfeature_names <- rownames(cfeatures_indices)[-1]

    feature_list[[i]] <- cfeature_names
    all_features <- c(all_features, unlist(cfeature_names))
}

sort(table(all_features), decreasing = TRUE)
```

# Robustness check of feature selection
## Univariate time-dependent cox
```{r}
sname <- paste0(".rds") # short l2so cv folds
lname <- paste0(".rds") # long l2so cv folds
data_sub_list <- readRDS(sname)
data_list <- readRDS(lname)
nfold <- length(data_sub_list)
```

```{r}
# List of variables to test
names(long_data)
x <- long_data[,10:72] # sensor features columns
variable_list <- names(x)
variable_list

features_selected <- list()
for (i in seq_along(data_list)) {
  d <- data_list[[i]]$train_data
 
# Perform univariate Cox regression for each variable and extract p-values
p_values <- sapply(variable_list, function(variable) {
  cox_model <- coxph(Surv(Time1, Time2, State) ~ get(variable), data = d, cluster=ID) # Index
  summary(cox_model)$coefficients[1, "Pr(>|z|)"]
})

# Set a significance level (e.g., 0.05) for filtering
significance_level <- 0.05

# Filter variables based on p-values
selected_variables <- variable_list[p_values < significance_level]

  features_selected[[i]] <- selected_variables
}

features <- unlist(features_selected)

library(dplyr)
sorted_df <- as.data.frame(table(features)) %>%
  arrange(desc(Freq))

sorted_df # frequency of each feature been selected acrosss folds
```
